---
title: "buildings"
output: html_document
---

This script creates two files:

* $parcels.Rdata$ contains parcels (polygon shapefile of tax assessor parcels within Sonoma County) and bldg_sonoma (polygon shapefile of building footprints within Sonoma County).
* $buildings.Rdata$ contains buildings, a dataframe with location, damage, & valuation information for residential buildings within the study area. 

Corinne Bowers

7/2/2021

```{r setup, include = FALSE}
## setup
knitr::opts_knit$set(root.dir = 'D:/1-PARRA/')

```

```{r message = FALSE}
## setup information
source('_data/setup.R')

## load area of interest (aoi)
load('_data/aoi/aoi.Rdata')

```

```{r parcels}
#### generate parcels.Rdata ####

# ## load buildings & parcels (takes a while)
# parcels <-
#   st_read('https://opendata.arcgis.com/datasets/2202c1cd6708441f987ca5552f2d9659_0.geojson', 
#           quiet = TRUE) %>%
#   st_transform(6417) %>% st_intersection(aoi)
# bldg_sonoma <- 
#   st_read(paste0('https://socogis.sonomacounty.ca.gov/map/rest/services/BASEPublic/Buildings/',
#                  'FeatureServer/0/query?outFields=*&where=1%3D1&f=geojson'), quiet = TRUE) %>%
#   st_transform(6417) %>% st_intersection(aoi)
# 
# ## save checkpoint
# save(parcels, bldg_sonoma, file = '_data/buildings/checkpoints/building_parcels.Rdata')

```

```{r buildings}
#### generate buildings.Rdata ####

# ## load buildings & parcels from file
# load('_data/buildings/checkpoints/building_parcels.Rdata')
# 
# ## load nonzero cells
# load('_data/nonzero/nonzero.Rdata')
# 
# ## crop buildings & parcels to only nonzero cells
# nonzero.buffer <- nonzero.buffer %>%
#   rasterToPolygons %>% 
#   st_as_sf %>% st_union %>% 
#   st_transform(6417)
# bldg_buffer <- bldg_sonoma[nonzero.buffer,] %>%
#   mutate(bldg_sqft = st_area(.)*mft^2) %>% 
#   st_centroid
# buildings <- st_intersection(bldg_buffer, parcels)
# 
# ## save checkpoint
# save(buildings, file = '_data/buildings/checkpoints/building_nonzero.Rdata')

```

```{r residential}
#### subset buildings dataframe to residential only ####

## load from file
load('_data/buildings/checkpoints/building_nonzero.Rdata')

## identify residential buildings
residential <- c(
  "ATTACHED UNIT",
  "CONDOMINIUM UNIT",
  "DETACHED UNIT IN A PUD",
  "DUET",
  "ENFORCEABLY RESTRICTED DWELLING",
  "MANUFACTURED HOME CONDOMINIUM LOT",
  "MANUFACTURED HOME ON URBAN LOT",
  "ONE DUPLEX (ONE STRUCTURE)",
  "RURAL RES SFD W/GRANNY UNIT",
  "RURAL RES W/MISC RES IMP",
  "RURAL RES/2 OR MORE RES",
  "RURAL RES/MANUFACTURED HOME",
  "RURAL RES/SECONDARY USE",
  "RURAL RES/SINGLE RES",
  "SFD SECONDARY USE",
  "SFD W/GRANNY UNIT",
  "SINGLE FAMILY DWELLING",
  "SINGLE LIVE/WORK UNIT",
  "TAXABLE MANUFACTURED HOME/CONDO LOT",
  "TAXABLE MANUFACTURED HOME/RENTED SITE",
  "TWO SFD ON SINGLE PARCEL",
  "WILDCAT SUBDIVISION LOT"
  )

## keep relevant columns & valuation codes
buildings <- buildings %>%
  filter(UseCType == 'Residential') %>%
  filter(UseCDesc %in% residential) %>%
  select("APN", "TxbltyDesc", "UseCDesc", "UseCType", "LndSzSF", "V601RollYr", "V601Land", 
         "V601Stru", "V601Fix", "V601FixRP", "V601Grow", "V601TotalL", "bldg_sqft") %>%
  mutate(bldg_sqft = as.numeric(bldg_sqft))

## filter out cases when a parcel has multiple structures
buildings <- buildings %>%
  cbind(st_coordinates(.)) %>%
  st_drop_geometry %>%
  group_by(APN) %>%
  summarize(across(everything(), ~.x[which.max(toNumber(bldg_sqft))])) %>%
  st_as_sf(coords = c('X', 'Y'), crs = 6417)

```

```{r geography}
#### attach geographic information ####

sonoma_tracts <- tracts(state = 'CA', county = 'Sonoma')
sonoma_cbgs <- block_groups(state = 'CA', county = 'Sonoma')
sonoma_blocks <- blocks(state = 'CA', county = 'Sonoma')
buildings <- buildings %>%
  st_intersection(sonoma_tracts %>% st_transform(6417) %>% transmute(tract = GEOID)) %>%
  st_intersection(sonoma_cbgs %>% st_transform(6417) %>% transmute(blockgroup = GEOID)) %>%
  st_intersection(sonoma_blocks %>% st_transform(6417) %>% transmute(block = GEOID10))

```

```{r resa}
#### attach safety tag information ####

## download RESA (Rapid Emergency Safety Assessment) data
# https://sonomacounty.maps.arcgis.com/apps/webappviewer/index.html?id=1cab5991f10643b1bc7c16e7769887c2
# https://sonomacounty.maps.arcgis.com/home/item.html?id=9e5d8762b5554765912f591b7540fed4
# instructions: open URL in ArcGIS Pro, then copy-paste Attributes table into Excel
resa <- read.csv('_data/buildings/resa.csv')
resa[resa == '<Null>'] <- NA

## attach RESA tag information
buildings <- buildings %>% 
  left_join(resa %>% transmute(id = OBJECTID, APN, RESA_Status_GIS, Within46ftInundation), 
            by = 'APN') %>% 
  mutate(status = case_when(RESA_Status_GIS != 'Orange' ~ RESA_Status_GIS) %>% 
           factor(levels = c('Red', 'Yellow', 'Green', 'N/A')))

```

```{r valuation}
#### attach information on valuation uncertainty ####

## find ACS median value of owner-occupied units by census tract
acs <- 
  getCensus(name = "acs/acs5", vintage = 2018, region = "tract:*", 
            regionin = "state:06+county:097", vars = "group(B25077)") %>%
  mutate(GEOID = paste0(state, county, tract) %>% toNumber) %>% 
  select(-c(GEO_ID, state, county, tract, NAME)) %>% 
  transmute(GEOID, acs.median = B25077_001E, MOE = B25077_001M) %>% 
  mutate(acs.05 = acs.median-MOE, acs.95 = acs.median+MOE) %>% 
  filter(acs.median > 0)

```

```{r valuation}
## plot ACS vs. Sonoma tax assessor roll values
buildings <- buildings %>% filter(V601TotalL <= 1e7)
buildings.plot <- buildings %>% 
  st_drop_geometry %>% 
  group_by(tract = toNumber(tract)) %>% 
  summarize(parcel.median = median(V601TotalL), 
            parcel.05 = quantile(V601TotalL, 0.05), 
            parcel.95 = quantile(V601TotalL, 0.95),
            n = length(V601TotalL)) %>% 
  left_join(acs %>% transmute(tract = toNumber(GEOID), acs.median, acs.05, acs.95), by = 'tract')
  
ggplot(buildings.plot) + 
  geom_segment(aes(x = acs.median, xend = acs.median, 
                   y = parcel.05, yend = parcel.95), color = 'grey70') + 
  geom_segment(aes(x = acs.05, xend = acs.95, 
                   y = parcel.median, yend = parcel.median), color = 'grey70') + 
  geom_point(aes(x = acs.median, y = parcel.median)) + 
  scale_x_origin('ACS Owner-Occupied Units (B25077)', 
                 labels = comma_format(scale = 1e-6, prefix = '$', suffix = 'M', accuracy = 0.1)) + 
  scale_y_origin('Sonoma Tax Assessor Roll', 
                 labels = comma_format(scale = 1e-6, prefix = '$', suffix = 'M', accuracy = 0.1)) + 
  ggtitle('Median SFH Valuation by Tract', subtitle = 'Sonoma County, 2019') + 
  geom_parity() + coord_fixed()

## determine the appropriate scaling factor
coef <- buildings.plot %>% 
  # filter(acs.median<=1e6 & parcel.median<=1e6) %>% 
  lm(parcel.median ~ acs.median+0, data = .) %>% 
  .$coefficients
buildings$value <- buildings$V601TotalL/coef

```

```{r valuation}
## plot ACS vs. Sonoma tax assessor roll uncertainty
ggplot(buildings.plot) + 
  geom_point(aes(x = acs.95-acs.05, y = (parcel.95-parcel.05)/sqrt(n))) + 
  scale_x_origin('ACS Owner-Occupied Units (B25077)', 
                 labels = comma_format(scale = 1e-6, prefix = '$', suffix = 'M', accuracy = 0.1)) + 
  scale_y_origin('Sonoma Tax Assessor Roll', 
                 labels = comma_format(scale = 1e-6, prefix = '$', suffix = 'M', accuracy = 0.1)) + 
  ggtitle('Valuation Margin of Error') + 
  geom_parity() + coord_fixed()
## note: here we are looking at the standard deviation of the population estimate, not individual houses

## attach valuation + uncertainty information
buildings <- buildings %>% 
  mutate(tract = toNumber(tract)) %>% 
  group_by(tract) %>% mutate(n = length(V601TotalL)) %>% ungroup %>% 
  left_join(acs %>% transmute(
    tract = toNumber(GEOID), 
    # acs.value = acs.median, 
    acs.sd = MOE/qnorm(0.95)*sqrt(n)), by = 'tract') %>% 
  select(-n)

```

```{r}
## renumber
buildings <- buildings %>% 
  arrange(APN) %>% 
  mutate(bldg = 1:nrow(.))

## save out
save(buildings, file = '_data/buildings/buildings.Rdata')

```

